s.function

### 2022-04-13 src/s_oa.Rmd
update

---
title: "diffusion, websites with 1 pc"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

@@

library(tidyverse)
library(rprojroot)
library(rgdal)
library(rgeos)
library(spdep)
library(tmap)
library(maptools)
library(knitr)
library(REAT)
library(tidygeocoder)
library(geosphere)

@@

@@

df <- bind_rows(df9610, df1112) %>% 
  filter(
    V1.domain < m,             # for multiple pc
    #V1.domain == n,             # for n == 1
    year > 1995 & year <2013) %>%   
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, oa11cd) %>%
  summarise(n = n()) %>%            # for multiple pc  
  #summarise(n = sum(V1.domain)) %>% # for n == 1 
  ungroup()

@@

## spatial data

@@

## lisa maps

@@

#path.out <- paste0(path, "/outputs/lisa_level_pc", n, "_oa") # for unique PC
path.out <- paste0(path, "/outputs/lisa_level_pc", m, "_oa")  # for multiple PC

morani <- data.frame()

#spatial weight matrix
sf_use_s2(FALSE)
nb <- poly2nb(oa.gb)
listw <- nb2listw(nb, style = "W", zero.policy = T)

for (t in 1996:2012){
test <- oa.gb
test@data <- test@data %>% left_join(df[df$year==t,], by = c("id"="oa11cd")) %>% 
  mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
         year = ifelse(is.na(year), t, year))

#spatial weight matrix
#moved out of the loop

#Moran's I
globalMoran <- moran.test(test$n, listw, zero.policy = T, na.action = na.exclude)
a <- cbind(globalMoran$estimate[[1]],globalMoran$p.value, t[1])
morani <- rbind(a, morani)

#LISA
lmoran<- cbind(test@data, localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude))
#lmoran

# centers the local Moran's around the mean
lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
lmoran$lag.n<-  lag.listw(listw,lmoran$n, NAOK = TRUE)

# centers the variable of interest around its mean

lmoran$ns <- lmoran$n - mean(lmoran$n, na.rm = TRUE) 
lmoran$lag.n <- lmoran$lag.n - mean(lmoran$lag.n, na.rm = TRUE)

signif <- 0.05
#lmoran

lmoran <- lmoran%>% 
  mutate(quadrant= ifelse(ns>0 & lag.n > 0, 1, 0)) %>% 
  mutate(quadrant= ifelse(ns<0 & lag.n < 0, 2, quadrant)) %>% 
  mutate(quadrant= ifelse(ns<0 & lag.n > 0, 3, quadrant)) %>% 
  mutate(quadrant= ifelse(ns>0 & lag.n < 0, 4, quadrant)) %>%   
  #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
  mutate(quadrant= ifelse(lmoran$`Pr(z != E(Ii))` > signif, 0, quadrant)) #%>%

#   mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st > 0, 1, 0)) %>% 
# mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st < 0, 2, quadrant2)) %>% 
#   mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st > 0, 3, quadrant2)) %>% 
#  mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st < 0, 4, quadrant2)) %>% 
#  mutate(quadrant2= ifelse(lmoran$LISA_PANC > signif, 0, quadrant2))

mun_merge_new<- merge(test, lmoran, by="id")

# R p value map
breaks = c(0, 1, 2, 3, 4, 5)

cities <- maps::world.cities %>% 
  filter(country.etc=="UK") %>% 
  arrange(pop) %>% tail(10) 
tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)

map <- tm_shape(mun_merge_new) + tm_fill(col = "quadrant", breaks = breaks, 
                                         palette = c("white","red","blue",
                                                     rgb(0,0,1,alpha=0.4),
                                                     rgb(1,0,0,alpha=0.4)), 
                                         labels = c("Not significant", "High-High","Low-Low","Low-High","High-Low"), 
                                         title="LISA for level, OA") +
  tm_legend(text.size = 1) +
  #tm_scale_bar(position = c("LEFT", "BOTTOM"),text.size = 1.0)+
  #tm_compass(type = "8star",   position = c("RIGHT", "BOTTOM"),      show.labels = 2,   text.size = 0.5)+
  #tm_borders(alpha=.1) +
  tm_borders(lwd = 0) +
  tm_shape(gb, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
  tm_shape(tif_sf) + tm_dots() + tm_text("name", size = .5) +
  tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ")
tmap_save(map, filename = paste(path.out, t, ".png", sep = ""))
}
```

Moran I

```{r}
morani %>% as_tibble() %>% 
  rename('morani'=V1,
                   p = V2,
                   year = V3) %>% 
  kable()

## Gini

gini.df <- data.frame()

for (t in 1996:2012){
    a <- df %>% filter(year==t)
  b <- as.data.frame(gini(a$n))
  b$year = t
  gini.df <- rbind(gini.df, b)
}
```

## Getis-Ord

t= 1998
local_g <- localG(OA.Census$Qualification, nb_lw)
local_g <- localG(df[df$year==t,]$n, listw)

local_g <- cbind(OA.Census, as.matrix(local_g))
names(local_g)[6] <- "gstat"

## Distances to cities and retail centres

@@

# Join with distance and area
df <- df %>% left_join(dist, by = c("oa11cd" = "oa11cd")) %>% 
  left_join(oa.gb@data, by = c("oa11cd" = "id")) %>% 
  relocate(area, .after = n) %>% 
  mutate(area = area / 1000000,
         density = n / area) %>% 
  relocate(density, .after = area)
sapply(df, function(x) sum(is.na(x)))
```

@@

# Join with complete panel
df <- df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd"))

sapply(df, function(x) sum(is.na(x)))
```

Website density regressions to London

```{r}
dens.London <- data.frame()
out.model <- data.frame()

for (t in 1996:2012){
  model <- lm(log(density+0.001) ~ log(London), data = df[df$year==t,])
  r2 <- as.data.frame(summary(model)$r.squared)
  r2$year = t
  dens.London <- rbind(r2, dens.London)

  out.model_ <- tidy(model, conf.int = TRUE) %>% 
    slice(2) %>% 
    mutate_if(is.numeric, round, 2)
  out.model <- rbind(out.model_, out.model)
}

names(dens.London)[1] <- "rsquared"
ggplot(dens.London) +
  geom_line(aes(x=year, y = rsquared))

library(broom)
out.model <- out.model %>% 
  mutate(year = 1996:2012)

ggplot(out.model, aes(x=reorder(year, year), y=estimate)) +
#ggplot(out.model, aes(x=year, y = estimate)) +  
           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
                       width = 0.2,size  = 1,
                       position = "dodge", color="turquoise4") +
  #geom_hline(yintercept = 0, color = "red", size = 1) +
  geom_point() + coord_flip() 
```

Website density regressions to the nearest city

```{r}
dens.distMet <- data.frame()
out.modelMet <- data.frame()

for (t in 1996:2012){
  model <- lm(log(density+0.001) ~ log(dist), data = df[df$year==t,])
  r2 <- as.data.frame(summary(model)$r.squared)
  r2$year = t
  dens.distMet <- rbind(r2, dens.distMet)

  out.model_ <- tidy(model, conf.int = TRUE) %>% 
    slice(2) %>% 
    mutate_if(is.numeric, round, 2)
  out.modelMet <- rbind(out.model_, out.modelMet)
}

names(dens.distMet)[1] <- "rsquared"
ggplot(dens.distMet) +
  geom_line(aes(x=year, y = rsquared))

out.modelMet <- out.modelMet %>% 
  mutate(year = 1996:2012)

ggplot(out.modelMet, aes(x=reorder(year, year), y=estimate)) +
           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
                       width = 0.2,size  = 1,
                       position = "dodge", color="turquoise4") +
  #geom_hline(yintercept = 0, color = "red", size = 1) +
  geom_point() + coord_flip() 
```

Website density regressions to the nearest retail centre

```{r}
dens.dist.retail <- data.frame()
out.model.retail <- data.frame()

for (t in 1996:2012){
  model <- lm(log(density+0.001) ~ log(dist.retail+0.001), data = df[df$year==t,])
  r2 <- as.data.frame(summary(model)$r.squared)
  r2$year = t
  dens.dist.retail <- rbind(r2, dens.dist.retail)

  out.model_ <- tidy(model, conf.int = TRUE) %>% 
    slice(2) %>% 
    mutate_if(is.numeric, round, 2)
  out.model.retail <- rbind(out.model_, out.model.retail)
}

names(dens.dist.retail)[1] <- "rsquared"
ggplot(dens.dist.retail) +
  geom_line(aes(x=year, y = rsquared))

out.model.retail <- out.model.retail %>% 
  mutate(year = 1996:2012)

ggplot(out.model.retail, aes(x=reorder(year, year), y=estimate)) +
           geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
                       width = 0.2,size  = 1,
                       position = "dodge", color="turquoise4") +
  #geom_hline(yintercept = 0, color = "red", size = 1) +
  geom_point() + coord_flip()

# library(pscl)
# m1 <- zeroinfl((n) ~ log(dist.retail+0.001),
#   data = df[df$year==t,], dist = "negbin", EM = TRUE)
# summary(m1)
# 
# df %>% filter(year == 1996) %>% 
#   group_by(group = cut(n, breaks = c(0, 1, 2, 10, 100, 1000, 10000), include.lowest = T, right = F)) %>%  #seq(0, max(n), 100))) %>%
#   summarise(n = sum(n))
```

Website density regressions to the nearest retail centre with spatial lag

```{r}
library(spatialreg)

model <- lagsarlm(log(density+0.1) ~ log(dist.retail+0.1), data = df[df$year==t,], 
                  listw, tol.solve = 1e-30, zero.policy = T)

dens.dist.retail <- data.frame()
for (t in 1996:2012){
  model <- lm(log(density) ~ log(dist.retail+0.1), data = df[df$year==t,])
  r2 <- as.data.frame(summary(model)$r.squared)
  r2$year = t
  dens.dist.retail <- rbind(r2, dens.dist.retail)
}
names(dens.dist.retail)[1] <- "rsquared"
ggplot(dens.dist.retail) +
  geom_line(aes(x=year, y = rsquared))

@@

## HH clusters

```{r}
hh <- mun_merge_new@data %>% filter(quadrant==1) %>% 
  select(id, quadrant) %>% 
  left_join(df[df$year==2000,], by =c("id" = "oa11cd")) %>% 
  select(id, year, n, area, density, London, dist, dist.retail)

hh.London <- lm(log(density+0.001) ~ log(London), data = hh)
summary(hh.London)
hh.dist <- lm(log(density+0.001) ~ log(dist), data = hh)
summary(hh.dist)
hh.retail <- lm(log(density+0.001) ~ log(dist.retail+0.001), data = hh)
summary(hh.retail)

df9610 <- read_csv(data.path9610) %>% 
  filter(V1.domain == n,
         year == 1999) %>% 
  left_join(lookup, by = "pc", suffix =c("",""))

hh %>% left_join(df9610, by =c("id"="oa11cd"))

Spatial regression

```{r}
library(splm)
m <- spml(log(n)~lag(log(n)), data = df, listw = listw, lag = TRUE, spatial.error = "b", 
          model = "within", effect = "individual", method = "eigen", na.action = na.fail)

>>>>

---
title: "s.function"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

@@

library(tidyverse)
library(broom)
library(rprojroot)

@@

## Load data

@@

df <- bind_rows(df9610, df1112) %>% 
  filter(
    #V1.domain < m,             # for multiple pc
    V1.domain == n,             # for n == 1
    year > 1995 & year <2013) %>%   
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, oa11cd) %>%
  #summarise(n = n()) %>%            # for multiple pc  
  summarise(n = sum(V1.domain)) %>% # for n == 1 
  ungroup()

@@

## Help, not really used

```{r, eval=FALSE}

df.test <- df %>% filter(oa11cd == "E00000007" |
                         oa11cd == "E00000021" |
                         oa11cd == "E00000025") %>% 
  mutate(rownum = row_number(),
         value2 = n/max(n))

# function to select unique groups
sample_n_groups = function(grouped_df, size, replace = FALSE, weight=NULL) {
  grp_var <- grouped_df %>% 
    groups %>%
    unlist %>% 
    as.character
  random_grp <- grouped_df %>% 
    summarise() %>% 
    sample_n(size, replace, weight) %>% 
    mutate(unique_id = 1:NROW(.))
  grouped_df %>% 
    right_join(random_grp, by=grp_var) %>% 
    group_by_(grp_var) 
}

df.test <- df %>% group_by(oa11cd) %>% 
  sample_n_groups(10000) %>% 
  arrange(oa11cd, year) %>% 
  select(-unique_id)

fit <- nls(n ~ SSlogis(year, Asym, xmid, scal), data = df.test)

fd.nls <- df.test %>% 
  group_by(oa11cd) %>%
  select(year, oa11cd, n) %>% 
  do(fitloop = tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = .))) %>% 
  unnest(fitloop)

fd.nls %>% filter(term=="xmid") %>% 
  summarise(rng = range(estimate))
```

## Test for S function estimation

To interpret the coefficients, check [SSlogis](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/SSlogis).
This is a self-starting model, which estimates the starting values.

```{r}
df.test <- df %>% filter(oa11cd == "S00103828")

fit <- nls(n ~ SSlogis(year, Asym, xmid, scal), data = df.test)
summary(fit)

df.test$prednls = predict(fit)

ggplot(df.test, aes(x = year, y = n) ) +
  geom_point() +
  geom_line(aes(y = prednls), size = 1) +
  scale_x_continuous("year", labels = as.character(df.test$year), breaks = df.test$year) +
  ylab("N. of websites")
  #scale_x_date(breaks = df.test$year) 
```

## S function estimation for all the country

```{r}
df.country <- df %>% group_by(year) %>% 
  summarise(n = sum(n))

fit <- nls(n ~ SSlogis(year, Asym, xmid, scal), data = df.country)
summary(fit)

df.country$prednls = predict(fit)

ggplot(df.country, aes(x = year, y = n) ) +
  geom_point() +
  geom_line(aes(y = prednls), size = 1) +
  scale_x_continuous("year", labels = as.character(df.country$year), breaks = df.country$year) +
  ylab("N. of websites") +
  ggtitle("Country")
  #scale_x_date(breaks = df.test$year) 
```

## Loop for S function estimation for OA

```{r}
a <- as_tibble()
all.oa <- unique(df$oa11cd)
for (oa in all.oa){
  data <- df %>% filter(oa11cd==oa) %>%
    select(year, oa11cd, n) 
    #lm(n~year, data = .)
    #tryCatch(tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = .)), , error=function(e) NULL)
  tryCatch(b <- tidy(nls(n ~ SSlogis(year, Asym, xmid, scal), data = data)), error=function(e) NULL)
  tryCatch(a <-rbind(a,b), error=function(e) NULL)
}

a %>% filter(term=="xmid") %>% 
  mutate(estimate = round(estimate, 0)) %>% 
  select(estimate) %>% 
  summarise(r = range(estimate))

a %>% filter(term=="xmid") %>% 
  mutate(estimate = round(estimate, 0)) %>% 
  filter(p.value < 0.01) %>%    # (227759 - 212782)/227759  # 6% not sig 
  select(estimate) %>% 
  ggplot(aes(x=estimate)) + 
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = 2003, colour = "red")
  scale_x_continuous("year", labels = round(a$estimate, 0), breaks = round(a$estimate, 0)) 
  #scale_x_discrete(labels = as.factor(a$estimate))

a <- a %>% filter(term=="xmid") %>% 
    bind_cols(all.oa) %>% 
    mutate(estimate = round(estimate, 0)) %>% 
    filter(p.value < 0.01) %>%    # (227759 - 212782)/227759  # 6% not sig 
    rename(id = ...6) %>% 
    select(id, estimate)
```

write the object with all the models

```{r}
path.out <- paste0(path, "/data/temp/s.csv")
write_csv(a, path.out)
```

## Map the xmid

Spatial data

@@

gb.df <- fortify(gb)
```

@@

# spatial transformations
oa.gb <- spTransform(oa.gb, CRS("+init=epsg:4326"))

# ggplot2 maps
oa.gb.f <- fortify(oa.gb, region = "id")

oa.gb.f <- oa.gb.f %>% left_join(a, by = "id") %>% 
  mutate(fast = ifelse(estimate < 2003, "fast", "slow")) %>% 
  arrange(order)  # if i don't order merge.nuts.f loses order and the map has gaps

# oa.gb.f %>% mutate(fast = ifelse(estimate < 2003, 1, 0)) %>% 
#   arrange(order) %>% 
ggplot(data = oa.gb.f %>% arrange(order), aes(x = long, y = lat)) +
  geom_polygon(aes(group = group, fill = estimate)) + #fill = fast, 
  #theme_nothing(legend = TRUE) +
  labs(title = "Clusters") +
  scale_fill_continuous(type = "viridis") +
  theme_void()
```

## OA classification

[Source](https://data.cdrc.ac.uk/dataset/output-area-classification-2011)

```{r}
oa.class.path <- paste0(path, "/data/raw/Output Area Classification")
oa.class <- readOGR(dsn = oa.class.path)
oa.class@data %>% glimpse()

pen.path <- paste0(path, "/data/raw/Output Area Classification/penportraits.csv")
pen <- read.csv(pen.path, sep = "\t")

oa.class.df <- a %>%  
  mutate(fast = ifelse(estimate < 2003, "fast", "slow")) %>% 
  left_join(oa.class@data, by = c("id"="OA_SA")) %>% 
  select(1:10) %>% 
  arrange(id)

length(unique(oa.class.df$GRP))

oa.class.df %>% 
  filter(fast == "slow") %>% 
  group_by(SUBGRP) %>% 
  summarise(n=n()) %>% 
  arrange(-n) %>% 
  mutate(freq = n / sum(n),
         freq = round(freq, 2),
         cum.n = cumsum(n),
         cum.freq = cumsum(n / sum(n)),
         cum.freq = round(cum.freq, 2)) %>% 
  left_join(pen %>% select(code, name),
            by = c("SUBGRP" = "code")) %>% 
  kable()

# no obvious group / subgroup patterns between slow and fast
```

## Regress xmid

### Distances to cities and retail centres

@@

# Join with distance and area
oa.class.df <- oa.class.df %>% left_join(dist, by = c("id" = "oa11cd")) 
sapply(df, function(x) sum(is.na(x)))
```

@@

# Join with complete panel
oa.class.df <- oa.class.df %>% left_join(dist.retail, by = c("id" = "oa11cd"))

sapply(oa.class.df, function(x) sum(is.na(x)))
```

Distance regressions

oa.class.df <- oa.class.df %>% left_join(df %>% group_by(oa11cd) %>% summarise(max.n = max(n, na.rm = T)), #filter(year==2012), 
                          by = c("id"="oa11cd"))

model.n <- lm(estimate ~ max.n, data = oa.class.df) #%>%filter(estimate>1996 & estimate <2012)
summary(model.n)
nobs(model.n)

model <- lm(estimate ~ log(London + 1) + log(dist + 1) + log(dist.retail + 1) + 
              as.factor(SPRGRP) + log(POPULATION) + max.n, data = oa.class.df) #%>%filter(estimate>1996 & estimate <2012)
summary(model)
nobs(model)

oa.class.df <- oa.class.df %>% mutate(fast.f = recode(fast, "fast" = 1, 
                                          "slow" = 0)) %>% 
  relocate(fast.f, .after = fast)

model.glm <- glm(fast.f ~ log(London + 1) + log(dist + 1) + log(dist.retail + 1) + 
              as.factor(SPRGRP) + log(POPULATION) + max.n, family = binomial(link='logit'), data = oa.class.df) #%>%filter(estimate>1996 & estimate <2012)
summary(model.glm)
nobs(model.glm)

library(pscl)
pR2(model.glm)

@@
